Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 Mw=7.9 Wenchuan earthquake

### 2014-09-13 paper.tex
Added remaining figures. Some text.

\documentclass{article}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=7in]{figs/lms_map_all.pdf}
    \caption{A map of the Longmen Shan region with the basemap showing shaded relief. The focal mechanism plot is located at the epicenter of the Wenchuan earthquake, while the thicker red line shows the surface trace of rupture. The thin red lines show 1km depth contours of our model geometry. GPS velocity vectors are colored black if they are not included in our parameter estimation and red if they are included. The blue line shows the two-dimensional cross-section we study. Thin black lines indicate other significant faults in the region.}
    \label{fig:regional_map}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{figs/stack_figure_all_details.pdf}
    \caption{a) Observed profile-parallel horizontal observed velocities are shown in black with 68\% confidence errorbars. Model-predicted velocities are shown by the red line with the surrounding gray region showing a 68\% confidence interval. b) Model geometry with colors and width along the locked fault showing predicted slip-deficit rates. Because shortening is the horizontal component of slip-deficit, increasing dips near the surface lead to great slip-deficits. c) Predicted horizontal velocity as a function of depth. Note the steep velocity gradient and corresponding strain accumulation above the end of the detachment surface.}
    \label{fig:big_stack}
\end{figure}
In

A Beichuan-only fault geometry would predict very little deformation beyond 100km (CITE savage 1983), leading one to exclude velocities beyond this range. Both the observed and predicted profile-parallel velocities shown in Figure \ref{fig:big_stack} make clear that such an estimate would result in a near-zero shortening estimate.

As demonstrated by the leveling-off of the model predicted velocities, velocities deeper into the plateau are likely controlled by other fault systems.

The debate over shortening in the Longmen Shan has centered around the clear lack of velocity gradient at the range front. 
Chen et al. 2000 < 3mm/yr of shortening in the Longmen Shan.
However, all their stations are within 50km of the range front. CHECK THIS
Reference the plot of GPS velocities in the figure which shows the best fitting solution.
General theme of previous interpretations: either ignored earthquake cycle effects or only looked at the near-field GPS.
Which, if there is a deep detachment, are useless because they would indicate almost no shortening.
Observations may also have been diminished because they were late in the earthquake cycle – just before Wenchuan.
Interseismic locking leads to a smooth profile of geodetic velocities across a locked fault.
Interpreting this profile is highly dependent on fault geometry.
In the Longmen Shan, a deep detachment leads to a surface velocity gradient that is localized very far to the west of the range front.
However, ignoring the detachment would indicate minimal shortening across the range front (see figures).

@@

>>>>

\documentclass[12pt]{article}

@@

\begin{figure}[h!]
    \centering
    \includegraphics{figs/lms_map_all.pdf}
    \caption{A map of the Longmen Shan region with the basemap showing shaded relief. The focal mechanism plot is located at the epicenter of the Wenchuan earthquake, while the thicker red line shows the surface trace of rupture. The thin red lines show 1km depth contours of our model geometry. GPS velocity vectors are colored black if they are not included in our parameter estimation and red if they are included. The blue line shows the two-dimensional cross-section we study. Thin black lines indicate other significant faults in the region.}
    \label{fig:regional_map}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics{figs/stack_figure_all_details.pdf}
    \caption{a) Observed profile-parallel horizontal observed velocities are shown in black with 68\% confidence errorbars. Model-predicted velocities are shown by the red line with the surrounding gray region showing a 68\% confidence interval. b) Model geometry with colors and width along the locked fault showing predicted slip-deficit rates. Because shortening is the horizontal component of slip-deficit, increasing dips near the surface lead to larger slip-deficits. c) Predicted horizontal velocity as a function of depth. Note the steep velocity gradient and corresponding strain accumulation above the end of the detachment surface.}
    \label{fig:big_stack}
\end{figure}

A Beichuan-only fault geometry would predict very little deformation beyond 100km (CITE savage 1983), leading one to exclude velocities beyond this range. Both the observed and predicted profile-parallel velocities shown in Figure \ref{fig:big_stack} make clear that such an estimate would result in a near-zero shortening estimate. A locked detachment, however, drives the predicted location of steep velocity gradient towards the northwest. As demonstrated by the leveling-off of the model predicted velocities, velocities deeper into the plateau are likely controlled by other fault systems.

The debate over shortening in the Longmen Shan has centered around the clear lack of velocity gradient at the range front. 
Chen et al. 2000 < 3mm/yr of shortening in the Longmen Shan.
However, all their stations are within 50km of the range front. CHECK THIS
Reference the plot of GPS velocities in the figure which shows the best fitting solution.
General theme of previous interpretations: either ignored earthquake cycle effects or only looked at the near-field GPS.
Which, if there is a deep detachment, are useless because they would indicate almost no shortening.
Observations may also have been diminished because they were late in the earthquake cycle – just before Wenchuan.
Interseismic locking leads to a smooth profile of geodetic velocities across a locked fault.
Interpreting this profile is highly dependent on fault geometry.
In the Longmen Shan, a deep detachment leads to a surface velocity gradient that is localized very far to the west of the range front.

\begin{figure}[h!]
    \centering
    \includegraphics{figs/depth_slip_contour.pdf}
    \caption{}
    \label{fig:distribution}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics{figs/hazard_all_details.pdf}
    \caption{Recurrence as a function of slip, assuming the best-fit slip rate for the Beichuan fault. The discrepancy between the Wenchuan-like recurrence and the paleoseismic estimate may be due to the presence of multiple active faults in the range-front.}
    \label{fig:hazard}
\end{figure}
